Optimize sequencing depth required for reliable profiling of microbiome by rarefaction and multimodel inference, using shotgun metagenomic data from pollination system and a modular profiling pipeline

### 2021-08-23 MResProject/Latex/Draft.tex
CMEECourseWork

\usepackage{setspace}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=red,citecolor=black]{hyperref}
\usepackage{lineno}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{floatrow}
\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{geometry}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{natbibspacing}
\usepackage[toc,page]{appendix}

\title{Metagenomic DNA discoveries after sequencing everything (bacteria, parasites, food, gut) inside a bee}
\author{Cong Liu}
\date{2021.8}

@@

  \maketitle
  \vspace*{\fill}
  \begin{center}
    A thesis submitted in partial fulfilment of the requirements for the degree of Master
of Science/Research at Imperial College London
\newline
Submitted for the MRes/MSc in Computational Methods in Ecology and Evolution
  \end{center}

  \newpage
  Declaration:
  \newline

  \newpage
  \linenumbers
  \begin{abstract}
    Bee-associated microbiome plays a crucial role in host health and relevant investigations have been boosted by shotgun metagenomics. 
    However, bioinformatics for metagenomic data analysis is complex, and the results may be compromised by insufficient sequencing. 
    Here, an integrated pipeline combining assembly-dependent and -independent methods was introduced for taxonomic and functional profiling of shotgun metagenomic data. 
    It was used for profiling metagenomic datasets ($2 \times 150 bp$ read pairs) from honey bees, bumble bees and flower surface. 
    Rarefaction and model fitting was conducted to compute expected species/function diversity (Hill numbers of order 0, 1, 2) returned by given sequencing effort in order to determine minimal sequencing depth capable of estimating diversity sufficiently. 
    As a result, functional profiling was demanding about sequencing effort. 
    None of datasets involved was sufficient for providing almost complete coverage of function diversity. 
    As for species diversity, the average minimal sequencing depth for reliable estimation of Hill number of order 0, 1 and 2 for honey bee samples are 40.33 million (12.10 Gbp), 18.57 million (5.57 Gbp) and 17.45 million (5.24 Gbp), respectively; and for bumble bee samples, the averages are 42.49 million (12.75 Gbp), 40.33 million (12.10 Gbp) and 24.77 (7.43 Gbp) million. 
    For flower surface, more than 0.88 million (0.26 Gbp) sequencing depth is needed. 
    These results are helpful for cost-effective utilization of shotgun metagenomics in investigations of pollination system. 
    The integrated pipeline and the framework for rarefaction analysis can also be applied for researches on other environments. 
  \end{abstract}

@@

  The health of bees is a crucial issue that has received increasing concern \citep{amiri2020using}. 
  As pollinators of numerous plants, bees play an important role in the stability of natural and agricultural systems \citep{hristov2020significance,foldesi2021larger,bansch2021crop,khalifa2021overview}. 
  However, diverse pathogens and environmental factors impose threaten on bee health. 
  The trend of bee decline has been widely observed at a global scale and attributed to multiple factors including pesticide exposure, invasive species, habitat loss, disease prevalence and climate change \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}.

@@

  Bee-associated microbiome plays a role in food digestion and carbohydrate metabolism by mediating processes including breaking down 
  % plant-produced polysaccharide including 
  plant-produced pectin and hemicellulose \citep{zheng2019division}; hydrolysis of sucrose \citep{engel2012functional,lee2015saccharide}, a predominant component of nectar \citep{nicolson2007nectar} and metabolism of mannose \citep{engel2012functional,lee2015saccharide} which presents in nectar 
  % of \textit{Tilia} 
  \citep{adler2000ecological} and is poisonous to honey bees \citep{de1986mechanism}. 
  Besides, it is also involved in immunity against pathogens, providing protection against pathogens including \textit{Crithidia}, 

@@

  \citep{ebeling2016biology,forsgren2010novel} and \textit{Nosema} 

@@

@@

  Host-associated microbiome plays an important role in host homeostasis and metagenomics boosted by high throughput sequencing provides powerful tools for understanding its composition, function and dynamics. 
  For example, amplicon sequencing is a powerful tool and has revealed incredible understanding in microbiome \citep{eckburg2005diversity,galloway2020tools,new2020metagenomics}. 
  However, it only focuses on a small part of the whole microbiome. 
  Shotgun sequencing of microbiome, or metagenomics, unselectively captures DNA in a sample and is capable of providing comprehensive inventories of taxa and functional gene clusters (FGCs) with high resolution \citep{quince2017shotgun,new2020metagenomics}.

@@

@@

  Typical goal of metagenomics is to provide taxonomic and functional profile of the community, and there is not a golden standard for bioinformatics of metagenome. 
  Generaly, one of the first steps in metagenomic analysis is assembling short reads into long contigs, which can help improve accuracy of metagenomic annotation \citep{tran2020assembling} and is necessary for 

@@

  However, metagenome assembly is complex, compromised by fragmental assembly, chimeras \citep{mikheenko2016metaquast} and loss of taxon/function diversity due to unassembled reads \citep{vollmers2017comparing,ayling2020new}. 
  Probably because of these shortcomings, assembly is skipped in some researches and short reads are directly proceeded for annotation \citep{tringe2005comparative,huson2007megan,abubucker2012metabolic,vermote2018amplicon,bovo2018shotgun}, although the accuracy may be compromised due to low information load of short reads \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling}. 
  A combination of both assembly-dependent and -free method might be helpful for overcoming the complexity of metagenomic assembly and inaccuracy of short read annotation \citep{becker2020modular}. 

@@

  Another challenge of metagenomics is the determination of sequencing depth. 
  It is recommended to retrieve as many reads as possible \citep{quince2017shotgun}, since insufficient sequencing causes imprecise estimation of taxon/FGC diversity \citep{cattonaro2018you,zaheer2018impact,pereira2019impact,gweon2019impact}. 
  However, deep sequencing is expensive, which hinders its utilization, especially in large-scale projects. 
  Currently, there is hardly any published guideline for the sufficient sequencing depth of a given environment or study type in order to reach a trade-off between sequencing effort and reliable output. 

@@

  The determination of sequencing depth can be conducted by rarefaction analysis, a method originated from traditional ecology \citep{sanders1968marine,hurlbert1971nonconcept,heck1975explicit,moreno2000assessing,hortal2005ed,gomez2014using,hughes2021sampling}. 
  In field based surveys, a random sample of individuals is drawn from a community and assigned to species. 
  Then a rarefaction curve illustrating expected biodiversity represented by given sampling effort (often measured by number of captured individuals) is generated by random subsampling the original sample without replacement, and quantified by model fitting \citep{hughes2005application,gotelli2011estimating}. 
  The slope of the rarefaction curve represents the expected rise of the curve if one more individual is captured. 
  The sample is nearly complete if and only if the final slope is small, and the point at which the slope of rarefaction curve falls to a cut-off value represents the minimal sampling effort required for assessment of biodiversity \citep{heck1975explicit,moreno2000assessing,hortal2004butterfly,hortal2005ed,chao2012coverage,gomez2014using,roswell2021conceptual}. 
  This framework for determination of sampling effort can be used in shotgun metagenomics, since a metagenomic dataset can be viewed as a random sample of an assemblage of genomic sequences, and 
  profiling is the process by which reads are assigned to taxa or FGCs. 
  The concept corresponding to sampling effort is sequencing depth in metagenomics, which is measured by read number.

@@

  In macroecology, comprehensive analytical frameworks are available for computing expected biodiversity returned by given sample sizes using a reference sample \citep{heck1975explicit,chao2012coverage,chao2014rarefaction}. 
  % However, analytical approaches derived for computing rarefaction curve in macroecology \citep{heck1975explicit,chao2012coverage,chao2014rarefaction} 
  However, they are not suitable for metagenomics because 
  % they are based on 
  of the assumptions that every captured individual can be assuredly assigned to a species, and the present of rare species is as reliable as that of abundant species. 
  In shotgun metagenomics, reads may be unannotated or annotated incorrectly due to low sequencing quality, host contamination, similar structures in genomes of different species, non-coding regions in genomes and limited sensitivity of profiling pipeline. 
  As a result, taxa or FGCs of low relative abundance are more likely to be false positive than that of high abundance.
  In this project, I aimed to (1) construct an integrated pipeline combining assembly-dependent and -independent methods for metagenomic profiling and (2) estimate minimal sequencing depth sufficient for covering species and FGC diversity of metagenome from three environmental types: honey bees, bumble bees and surface of flowers. 
  In the integrated pipeline (Figure \ref{Pipeline}), assembly-dependent taxon profiling is conducted after quality filtering and removing host contamination. 

@@

  This integrated pipeline was used to profile metagenomic datasets from honey bees, bumble bees and environmental DNA (eDNA) washed from a flower (\textit{Erigeron annuus}). 
  It was shown that the construction of reference database and assembly-free search helped identified species not represented by the assembly. 
  Then I simulated different sequencing depth by rarefaction and profiled subsampled datasets using the integrated pipeline, generating inventories of species and Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}. 
  Each KO represents one FGC, \textit{i.e.} an aggregate of genes with same function. 

@@

  The relationship between Hill numbers and sequencing depth (number of 150 bp pair-end reads) was quantified by fitting and averaging asymptotic species accumulation models, and estimation of minimal sequencing depth was given by the point where the slope of rarefaction curve drops to cut-off values. 
  Sample would be excluded from estimation of minimal sequencing depth if the original dataset was not sufficient for providing an almost complete inventory of species or KOs, which was characterized by small final slope of rarefaction curve for species or KO richness (Hill number of order 0). 
  As a result, when measuring species diversity by Hill number of order 0, 1 and 2, the average minimal sequencing depth for honey bee samples are 40.33 million, 18.57 million and 17.45 million; for bumble bee samples, the averages are 42.49 million, 40.33 million and 24.77 million; and for flower surface, 0.88 million sequencing depth is not sufficient for providing an almost complete inventory of species, indicating higher sequencing depth is needed. 
  As for functional diversity, no datasets involved in this project is big enough for providing an almost complete inventory of KOs, indicating the sequencing depth need to be higher than the maximum of each sample type, \textit{i.e.} 45.06 million for honey bees, 53.61 million for bumble bees and 0.88 million for flower surface.

@@

  \section{Materials and Methods}
    \subsection{Samples, DNA extraction and sequencing}

    \subsection{Metagenomic profiling using integrated pipeline}
      \subsubsection{Integrated pipeline}
      The integrated pipeline was designed for shotgun metagenomic profiling (assigning reads to taxa or FGCs). 
      It is separated into five modules (Figure \ref{Pipeline}).
      \newline
      In the module of quality control and host removing, sequencing data quality is checked using FastQC v.0.11.5 \citep{andrews2010fastqc} and quality control is conducted by Trimmomatic v.0.39 \citep{bolger2014trimmomatic}. 
      Then clean reads are mapped to host genome using Bowtie2 v.2.4.2 \citep{langmead2012fast} and non-host reads are extracted by SAMtools v.1.11 \citep{li2009sequence}. 
      \newline
      The non-host reads are subject to the module of assembly-dependent species identification. 
      \textit{De novo} assembly is conducted by SPAdes v.3.15.2 \citep{prjibelski2020using}. 
      Assembled contigs are aligned to NCBI non-redundant (nr) database by DIAMOND v.2.0.7.145 \citep{buchfink2015fast}, and assigned to taxa by MEGAN6 \citep{huson2007megan}. 
      \newline
      Then fragment recruitment is conducted. 
      A reference database comprising reference genome dataset, \textit{i.e.} genomic sequences in FASTA format and corresponding genome annotation in general feature format (gff), was constructed. 
      For each species represented by assembly, its reference genome dataset, if available, is downloaded from NCBI using its \textit{datasets} command-line tool and added to the reference database. 
      Then non-host reads are mapped to the reference database by Bowtie2, and unmapped reads are extracted by SAMtools. 
      \newline
      Reads not recruited by the reference database are subjected to assembly-independent species identification. 
      They are aligned to NCBI nr database through DIAMOND and assigned to taxa by MEGAN6.
      \newline
      Finally, functional annotation is conducted by EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      It takes coding sequences (CDSs) of genomes in the reference database and reads subject to assembly-independent species identification as input and assigns them to KOs.
      \newline
      The integrated pipeline was used for analyzing metagenomic datasets involved in this study and details in parameter settings of each module are described in Supplementary \ref{ParameterSetting}.

@@

    \subsection{Estimation of minimal sequencing depth required for metagenomic profiling}
    The impact of sequencing depth on taxon/function diversity of metagenome was simulated by rarefaction. 
    Diversity was measured by Hill numbers, which are calculated from inventory of relative abundances. 
    For taxon diversity, the inventory was generated by assigning reads to species; while for function diversity, it was obtained by assigning reads to KOs. 
    Then the rarefaction curve plotting Hill number against sequencing depth was computed, and minimal sequencing depth required for reliable extimation of diversity were given by the point where the slope of rarefaction curve drops to a cut-off value.

@@

      It provides an estimation of the total diversity and is comparable among metagenomic DNA samples with different sequencing depth \citep{lamas1991preliminary,soberonm1993use,hortal2004butterfly,jimenez2005determining,hortal2006evaluating}. 
      However, the accuracy of asymptotic estimators is controversial in macroecology \citep{soberonm1993use,colwell1994estimating,chazdon1998statistical,jimenez2006comparing,hortal2006evaluating}, and their performance in shotgun metagenomics is seldom evaluated.

@@

    % \textit{Candidatus Schmidhempelia bombi} is an uncultured symbiont of \textit{Bombus impatiens} \citep{martinson2014genomic}. 
    % It should be noted that some typical bee-associated bacteria were found on the flower, including \textit{Bartonella apis}, \textit{Bifidobacterium asteroides}, \textit{Bombilactobacillus mellis} and \textit{Gilliamella apicola}.

@@

    Figure \ref{ArthropodHeatmap} shows composition of arthropods, and most of them are pollinators within \textit{Apis} and \textit{Bombus}. 

@@

@@

    Figure \ref{PlantHeatmap} illustrates plant species, indicating foraging area of bees. 
    Most samples were dominanted by one or two plant species, except \textit{Bee\_Amellifera\_hv13\_2}. 
    Some crops were identified, including \textit{Brassica napus} (rape), \textit{Brassica oleracea}, \textit{Cicer arietinum} (chickpea), \textit{Glycine max} (soybean), \textit{Helianthus annuus} (sunflower), \textit{Nicotiana sylvestris} (flowering tabacco) and \textit{Raphanus sativus} (radish).
    Figure \ref{FungusHeatmap} shows compositions of fungal species. 
    \textit{Nosema ceranae}, a widespread bee pathogen, was the dominant fungal species in most samples including the flower eDNA. 
    Three yeast species (\textit{Clavispora lusitaniae}, \textit{Saprochaete ingens} and \textit{Wickerhamiella sorbophila}) were found in bees. 
    % Other fungal species include three yeasts (\textit{Clavispora lusitaniae}, \textit{Saprochaete ingens} and \textit{Wickerhamiella sorbophila}) 

@@

    Figure \ref{ViruseHeatmap} illustrates viruses identified. 

>>>>

\usepackage{setspace}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=red,citecolor=black]{hyperref}
\usepackage{lineno}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{floatrow}
\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{geometry}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{natbibspacing}
\usepackage[toc,page]{appendix}
\usepackage{makecell}

% Metagenomic DNA discoveries after sequencing everything (bacteria, parasites, food, gut) inside a bee
\title{Optimize sequencing depth required for reliable profiling of microbiome by rarefaction and multimodel inference, using shotgun metagenomic data from pollination system and a modular profiling pipeline}
\author{Cong Liu}
\date{2021.8}

@@

  \maketitle
  \vspace*{\fill}
  \begin{center}
    A thesis submitted in partial fulfilment of the requirements for the degree of Master of Research at Imperial College London
\newline
Submitted for the MRes in Computational Methods in Ecology and Evolution
  \end{center}

  \newpage
  Declaration:
  \newline
  All datasets used in this project are provided by Dr. Peter Graystock and the project was conducted under his supervison. 
  I am responsible for the development and conduction of the analyses presented.

  \newpage
  \linenumbers
  \begin{abstract}
    Bee-associated microbiome plays a crucial role in host health and relevant investigations have been boosted by shotgun metagenomics, which provides unique advantages by capturing diversity of multiple taxonomic clades and providing information on function potentiality. 
    However, utilization of shotgun metagenomics is hindered by complexity of data analysis and high cost of sequencing. 
    % However, bioinformatics for metagenomic data analysis is complex, and the results may be compromised by insufficient sequencing depth. 
    Here, an integrated pipeline combining assembly-dependent and -independent methods was introduced for taxonomic and functional profiling of shotgun metagenomic data, and a framework of rarefaction and multimodel inference was constructed for optimizing sequencing depth. 
    Both the pipeline and the framework were used for analysis of real metagenomic datasets ($2 \times 150 bp$ read pairs) from honey bees, bumble bees and flower washes. 
    The integrated pipeline illustrated taxon composition and metabolic potentiality of metagenome, and provided improvement in species identification compared with an assembly-dependent procedure. 
    Sequencing depth can be optimized for species diversity estimation especially when little emphasis is given to rare species, while functional profiling requires deeper sequencing than taxon profiling. 
    % Rarefaction and model fitting was conducted to compute expected species/function diversity (Hill numbers of order 0, 1, 2) returned by given sequencing effort in order to optimize sequencing depth, balancing sequencing effort and capability of estimating diversity reliably. 
    % As a result, functional profiling was demanding about sequencing effort. 
    % None of datasets involved was sufficient for providing almost complete coverage of function diversity. 
    % As for species diversity, the average optimal sequencing depth for reliable estimation of Hill number of order 0, 1 and 2 for honey bee samples are 40.33 million (12.10 Gbp), 18.57 million (5.57 Gbp) and 17.45 million (5.24 Gbp), respectively; and for bumble bee samples, the averages are 42.49 million (12.75 Gbp), 40.33 million (12.10 Gbp) and 24.77 (7.43 Gbp) million. 
    % For flower surface, more than 0.88 million (0.26 Gbp) sequencing depth is needed. 
    These results are helpful for cost-effective utilization of shotgun metagenomics in investigations of pollination system. 
    % The integrated pipeline and the framework for rarefaction analysis can also be applied for researches on other environments. 
  \end{abstract}

@@

  Pollinators such as bees are crucial to maintain global food security and provide stability of natural systems \citep{hristov2020significance,bansch2021crop,khalifa2021overview} but their health is under increasing concern \citep{amiri2020using}. 
  % The health of bees is a crucial issue that has received increasing concern \citep{amiri2020using}. 
  % As pollinators of numerous plants, bees play an important role in the stability of natural and agricultural systems \citep{hristov2020significance,bansch2021crop,khalifa2021overview}.
  % foldesi2021larger 
  Diverse pathogens and environmental factors threaten bee health and are contributing to their declines. 
  % The trend of bee decline has been widely observed at a global scale and attributed to multiple factors including 
  A combination of stressors including parasites, pesticide exposure, invasive species, habitat loss and climate change are all contributing to declines of bee populations. \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}.

@@

  Within the gut of bees, stable microbial communities (microbiomes) play crucial roles in food digestion, parasite defence and chemical detoxification. 
  They mediate food digestion processes including 
  % plant-produced polysaccharide including pectin and hemicellulose \citep{zheng2019division}
  polysaccharides breakdown \citep{zheng2019division}, sucrose hydrolysis \citep{engel2012functional,lee2015saccharide}, 
  % a predominant component of nectar \citep{nicolson2007nectar} 
  and mannose metabolism \citep{engel2012functional,lee2015saccharide}. 
  % which presents in nectar of \textit{Tilia} \citep{adler2000ecological} and is poisonous to honey bees \citep{de1986mechanism}. 
  They also provide protection against pathogens including \textit{Crithidia}, 

@@

  \citep{ebeling2016biology,forsgren2010novel} and \textit{Nosema sp.} 

@@

  Besides, bee-associated symbionts can aid in detoxification. 
  They are involved in resistance to both metal and metalloid toxins including cadmium \citep{rothman2019cadmium}, copper \citep{rothman2020direct} and selenate \citep{rothman2019bumble}. 

@@

  Amplicon sequencing, in which a species-specific barcode region is amplified and sequenced, is a powerful method for investigations of microbiome \citep{abdelfattah2018metabarcoding}.  
  It is vastly used in bee microbiome investigations to illustrate the taxonomic diversity of bee-associated bacteria and fungi (\textit{e.g.} \cite{geldert2021dietary,wang2021gut,powell9field,kapheim2021composition}). 
  However, bees visit numerous niches during environment exploration and foraging activities, and can get contact with diverse eDNA signatures, which provide insight into bee ecology by reflecting interactions between bees and other organisms including bacteria, fungi, plants, arthropods and viruses \citep{bovo2018shotgun,ribani2020honey,bovo2020shotgun,matsuzawa2020application}. 
  It is difficult to explore this diversity via amplicon sequencing because it only captures a fraction of the whole community since analysis of different taxonomic groups is based on different barcode regions, \textit{e.g.} 16S ribosomal RNA (rRNA) for bacteria \citep{hayashi2002phylogenetic,eckburg2005diversity}, internal transcribed spacer (ITS) for fungi \citep{nilsson2008intraspecific}, cytochrome c oxidase subunit I (COI) for Animalia \citep{hebert2003barcoding} 
  As a result, amplicon sequencing only captures taxon diversity within a certain clade. 
  Besides, it is difficult to illustrate function potentiality of bee-associated microbiome using amplicon sequencing since it does not provide information on content of functional gene clusters (FGCs), \textit{i.e.} aggregates of genes with same function. 
  As a result, functional capacity needs to be inferred based on reference genomes \citep{asshauer2015tax4fun,douglas2018predicting}. 
  However, bee bacterial symbionts are diversified at strain level \citep{engel2012functional,powell2016strain,ellegaard2020vast} and bacterial strains are often highly variable in gene content \citep{cordero2014explaining,brockhurst2019ecology}. 
  As a result, amplicon sequencing-based inference of functional capacity of bee microbiome may not be reliable.
  % Host-associated microbiome plays an important role in host homeostasis and metagenomics boosted by high throughput sequencing provides powerful tools for understanding its composition, function and dynamics. 
  % For example, amplicon sequencing is a powerful tool and has revealed incredible understanding in microbiome \citep{eckburg2005diversity,galloway2020tools,new2020metagenomics}. 
  % However, it only focuses on a small part of the whole microbiome. 
  % Shotgun sequencing of microbiome, or metagenomics, unselectively captures DNA in a sample and is capable of providing comprehensive inventories of taxa and functional gene clusters (FGCs) with high resolution \citep{quince2017shotgun,new2020metagenomics}.

@@

  Shotgun metagenomics provides an alternative for microbiome investigations to overcome drawbacks of amplicon sequencing. 
  By capturing and sequencing DNA fragments unselectively, shotgun metagenomics is capable of providing comprehensive inventories of taxa and FGCs \citep{quince2017shotgun,new2020metagenomics,galloway2020tools}. 

@@

  Typical goal of metagenomics is to provide taxonomic and functional profile of microbiome, and there is not a golden standard for bioinformatics of metagenome. 
  Generally, one of the first steps in metagenomic analysis is assembling short reads into long contigs, which can help improve accuracy of metagenomic annotation \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling} and is necessary for 

@@

  However, metagenome assembly is complex, compromised by fragmental assembly, chimaeras \citep{mikheenko2016metaquast} and loss of taxon/function diversity due to unassembled reads \citep{vollmers2017comparing,ayling2020new}. 
  Probably because of these shortcomings, assembly is skipped in some researches and short reads are directly proceeded for annotation \citep{tringe2005comparative,abubucker2012metabolic,vermote2018amplicon,bovo2018shotgun}, although the accuracy can be compromised due to low information load of short reads \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling}. 
  A combination of both assembly-dependent and -free methods could overcome the complexity and improving accuracy of metagenomic profiling \citep{becker2020modular}.
  % complexity of metagenomic assembly and inaccuracy of short read annotation \citep{becker2020modular}. 

@@

  A remaining challenge of metagenomics is the determination of a sequencing depth that provides reliable estimation of taxon/FGC diversity without overspending. 
  It is recommended to retrieve as many reads as possible \citep{quince2017shotgun} since insufficient sequencing causes compromise in metagenome profiling \citep{cattonaro2018you,zaheer2018impact,pereira2019impact,gweon2019impact}. 
  However, deep metagenomic sequencing is expensive, which hinders its utilization, especially in large-scale projects. 
  Currently, there are few published guidelines for the sufficient sequencing depth of a given environment or study type in order to reach a trade-off between sequencing effort and reliable output. 

@@

  In order to balance sequencing cost and reliable estimation of taxon/FGC diversity, expected diversity represented by given sequencing depth need to be computed. 
  Since a metagenomic dataset can be viewed as a random sample of an assemblage of genomic sequences, and profiling is the process by which reads are assigned to taxa or FGCs, the relationship between sequencing depth and diversity can be illustrated by rarefaction (randomly subsampling the original dataset without replacement) and quantified by model fitting, if the original dataset can provide an almost complete inventory of taxa/FGCs \citep{heck1975explicit,hortal2005ed,gomez2014using,hughes2021sampling}. 
  The sequencing depth is sufficient for reliable estimation of taxon/FGC diversity if and only if the slope of the model is small \citep{hortal2005ed,chao2012coverage,roswell2021conceptual}. 
  % The determination of sequencing depth can be conducted by rarefaction analysis, a method widely used in traditional ecology \citep{heck1975explicit,hortal2005ed,gomez2014using,hughes2021sampling}. 
  % In field based surveys, a random sample of individuals is drawn from a community and assigned to species. 
  % Then a rarefaction curve illustrating expected biodiversity represented by given sampling effort (often measured by number of captured individuals) is generated by random subsampling the original sample without replacement, and quantified by model fitting \citep{hughes2005application,gotelli2011estimating}. 
  % The slope of the rarefaction curve represents the expected rise of the curve if one more individual is captured. 
  % The sample is nearly complete if and only if the final slope is small, and the point at which the slope of rarefaction curve falls to a cut-off value represents the minimal sampling effort required for assessment of biodiversity \citep{hortal2005ed,chao2012coverage,gomez2014using,roswell2021conceptual}. 
  % This framework for determination of sampling effort can be used in shotgun metagenomics, since a metagenomic dataset can be viewed as a random sample of an assemblage of genomic sequences, and 
  % profiling is the process by which reads are assigned to taxa or FGCs. 
  % The concept corresponding to sampling effort is sequencing depth in metagenomics, which is measured by read number.
  % In macroecology, comprehensive analytical frameworks are available for computing expected biodiversity returned by given sample sizes using a reference sample \citep{heck1975explicit,chao2012coverage,chao2014rarefaction}. 
  % % However, analytical approaches derived for computing rarefaction curve in macroecology \citep{heck1975explicit,chao2012coverage,chao2014rarefaction} 
  % However, they are not suitable for metagenomics because 
  % % they are based on 
  % of the assumptions that every captured individual can be assuredly assigned to a species, and the present of rare species is as reliable as that of abundant species. 
  % In shotgun metagenomics, reads may be unannotated or annotated incorrectly due to low sequencing quality, host contamination, similar structures in genomes of different species, non-coding regions in genomes and limited sensitivity of profiling pipeline. 
  % As a result, taxa or FGCs of low relative abundance are more likely to be false positive than that of high abundance.

@@

  In this project, I utilised metagenomes from three environmental types: the gut of a honey bee (\textit{Apis mellifera}), the gut of a common North American bumble bee (\textit{Bombus impatiens}) and the surface of a flower (\textit{Erigeron annuus}). 
  I aimed to (1) develop an integrated pipeline combining assembly-dependent and -independent methods to deliver improved taxon annotation of sequencing data and (2) optimize sequencing depth to balance sequencing cost and requirement for reliable analysis of microbial species diversity, or the description of their functional diversity. 
  In the integrated pipeline (Figure \ref{Pipeline}), 
  assembly-dependent taxon profiling is conducted after quality filtering and removing host contamination. 

@@

  This integrated pipeline was used to profile metagenomic datasets involved in this project. 
  % It was shown that the construction of reference database and assembly-free search helped identified species not represented by the assembly. 
  Then in order to estimate optimal sequencing depth for reliable representing taxonomic/functional diversity of metagenomes, I simulated different sequencing depth by rarefaction and profiled subsampled datasets using the integrated pipeline, generating inventories of species and FGCs represented by Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}. 

@@

  The relationship between Hill numbers and sequencing depth was quantified by fitting and averaging asymptotic species accumulation models, and estimation of optimal sequencing depth was given by the point where the slope of rarefaction curve drops to cut-off values. 
  % Sample would be excluded from optimization of sequencing depth if the original dataset was not sufficient for providing an almost complete inventory of species or KOs, which was characterized by small final slope of rarefaction curve for species or KO richness (Hill number of order 0). 
  % As a result, when measuring species diversity by Hill number of order 0, 1 and 2, the average optimal sequencing depth for honey bee samples are 40.33 million, 18.57 million and 17.45 million; for bumble bee samples, the averages are 42.49 million, 40.33 million and 24.77 million; and for flower surface, 0.88 million sequencing depth is not sufficient for providing an almost complete inventory of species, indicating higher sequencing depth is needed. 
  % As for functional diversity, no datasets involved in this project is big enough for providing an almost complete inventory of KOs, indicating the sequencing depth need to be higher than the maximum of each sample type, \textit{i.e.} 45.06 million for honey bees, 53.61 million for bumble bees and 0.88 million for flower surface. 
  % The integrated pipeline and the framework for rarefaction analysis can also be applied for investigations of microbiomes from other environments, and the estimations of optimal sequencing depth provide guidelines for cost-effective utilization of shotgun metagenomics in investigations of pollination system.

@@

  \section{Materials and Methods}
    \subsection{Samples, DNA extraction and sequencing}
    Samples include four honey bees (two from the same hive (hive 13), one from a hive in the same apiary (hive 15) and one caught foraging), three bumble bees from commercially supplied (Biobest) bumblebee colonies (two from the same colony), and one buffer wash of a wild flower (\textit{Erigeron annuus}).  
    \newline
    DNA extraction was performed as in \cite{graystock2020dominant}, followed by library preparation using a template protocol with Illumina Nextera Library Prep kits. Briefly, this involved tagmentation into fragments of 300 bases before eight samples of 10ng were pooled together and sent to Beijing Genomics Institute for further quality control and sequencing using a full lane in the X-ten platform.

    \subsection{Metagenomic profiling using integrated pipeline}
      \subsubsection{Integrated pipeline}
      The integrated pipeline was designed for shotgun metagenomic profiling, \textit{i.e.} assigning reads to taxa or FGCs. 
      It is separated into five modules (Figure \ref{Pipeline}).
      \newline
      In quality control and host removing, raw sequencing data quality is checked using FastQC v.0.11.5 \citep{andrews2010fastqc} and filtered by Trimmomatic v.0.39 \citep{bolger2014trimmomatic}. 
      Then clean reads are mapped to host genome using Bowtie2 v.2.4.2 \citep{langmead2012fast} and non-host reads are extracted by SAMtools v.1.11 \citep{li2009sequence}. 
      \newline
      The non-host reads are subject to the module of assembly-dependent species identification. 
      \textit{De novo} assembly is conducted by SPAdes v.3.15.2 \citep{prjibelski2020using}. 
      Assembled contigs are aligned to NCBI non-redundant (nr) database by DIAMOND v.2.0.7.145 \citep{buchfink2015fast}, and assigned to taxa by MEGAN6 \citep{huson2007megan}. 
      \newline
      Then fragment recruitment is conducted. 
      A reference database is constructed and it comprises reference genome dataset, \textit{i.e.} genomic sequences in FASTA format and corresponding genome annotation in general feature format (gff). 
      For each species represented by assembly, its reference genome dataset, if available, is downloaded from NCBI using its \textit{datasets} command-line tool and added to the reference database. 
      Then non-host reads are mapped to the reference database by Bowtie2, and unmapped reads are extracted by SAMtools. 
      \newline
      Reads not recruited by the reference database are subjected to assembly-independent species identification. 
      They are aligned to NCBI nr database through DIAMOND and assigned to taxa by MEGAN6.
      \newline
      Finally, functional annotation is conducted by EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      It takes coding sequences (CDSs) of genomes in the reference database and reads subject to assembly-independent species identification as input and assigns them to KOs.
      \newline
      The integrated pipeline was used for analyzing metagenomic datasets involved in this study and details in parameter settings of each module are described in Supplementary \ref{ParameterSetting}.

@@

    \subsection{Estimation of optimal sequencing depth required for metagenomic profiling}
    % The impact of sequencing depth on taxon/function diversity of metagenome was simulated by rarefaction. 
    % Diversity was measured by Hill numbers, which are calculated from inventory of relative abundances. 
    % For taxon diversity, the inventory was generated by assigning reads to species; while for function diversity, it was obtained by assigning reads to KOs. 
    % Then the rarefaction curve plotting Hill number against sequencing depth was computed, and optimal sequencing depth required for reliable estimation of diversity were given by the point where the slope of rarefaction curve drops to a cut-off value.

@@

      It provides an estimation of the total diversity and is comparable among metagenomic DNA samples with different sequencing depth. 
      % \citep{lamas1991preliminary,soberonm1993use,hortal2004butterfly,jimenez2005determining,hortal2006evaluating}. 
      However, the accuracy of asymptotic estimators is controversial \citep{colwell1994estimating,chazdon1998statistical,jimenez2006comparing,hortal2006evaluating}.

@@

@@

    Composition of arthropods and plants indicating interactions within pollination networks. 
    Most arthropods identified are pollinators within \textit{Apis} and \textit{Bombus} (Figure \ref{ArthropodHeatmap}). 

@@

    As for plant species (Figure \ref{PlantHeatmap}), they indicate foraging areas of bees. Several crop species were identified, including \textit{Brassica napus} (rape), \textit{Brassica oleracea}, \textit{Cicer arietinum} (chickpea), \textit{Glycine max} (soybean), \textit{Helianthus annuus} (sunflower), \textit{Nicotiana sylvestris} (flowering tabacco) and \textit{Raphanus sativus} (radish). 

@@

    Fungal and viruse species were identified. 
    In fungal communities of most samples, \textit{Nosema ceranae}, a widespread bee pathogen, was the dominant species. (Figure \ref{FungusHeatmap}). 
    Besides, three yeast species (\textit{Clavispora lusitaniae}, \textit{Saprochaete ingens} and \textit{Wickerhamiella sorbophila}) were also found in bees.  

@@

    As for viruses, most of them are phages or arthropod-infecting species (Figure \ref{ViruseHeatmap}). 
    Phage species include \textit{Bifidobacterium phage BitterVaud1} infecting bee-commensal bacterium \textit{Bifidobacterium asteroides} \citep{bonilla2020honey}; \textit{Bacteriophage sp.} infecting \textit{Pseudomonas aeruginosa} \citep{essoh2015investigation}, an opportunistic pathogen that might contaminate bees \citep{bailey1968honey,papadopoulou1992experimental,papadopoulou1993haemocyte}; and unclassified species within Myoviridae and Siphoviridae. 
